---
title: "DATA 607 - Assignment 1"
author: "Albert Gilharry"
date: "February 3, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r load-libraries}
library(tidyverse)
```

## Load Data
<i>The original dataset is located at https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data</i>
```{r load-data}
mushrooms <- read.csv('https://raw.githubusercontent.com/albert-gilharry/data607-assignment1/master/agaricus-lepiota.data',header = FALSE)
```

#Rename Columns
<i>Rename all columns for the sake of completeness.</i>
```{r rename-columns}
names(mushrooms) <- c("class","cap-shape","cap-surface","cap-color","bruises","odor","gill-attachment","gill-spacing","gill-size","gill-color","stalk-shape","stalk-root","stalk-surface-above-ring","stalk-surface-below-ring","stalk-color-above-ring","stalk-color-below-ring","veil-type","veil-color","ring-number","ring-type","spore-print-color","population","habitat")
```


## Extract Columns of Interest
```{r select-columns}
mushrooms <- select(mushrooms, "class", "cap-shape", "cap-color","odor","habitat")
```

## Preview Data
```{r preview-data}
head(mushrooms)
```
<i>Transformations are necessary to make more sense of this data.<i/>

## Examine Data Structure
```{r examine-structure}
str(mushrooms)
```

## Replace Abbreviations
<i>Careful considerations were taken to ensure the abbrevitions in the levels references the correct metadata in the dictionary.</i> 
```{r replace-abbreviations}
levels(mushrooms$class)
levels(mushrooms$class) <- c("edible","poisonous")

levels(mushrooms$`cap-shape`)
levels(mushrooms$`cap-shape`) <- c("bell","conical","flat","knobbed","sunken","convex")

levels(mushrooms$`cap-color`)
levels(mushrooms$`cap-color`) <- c("buff","cinnamon","red","gray","brown","pink","green","purple","white","yellow")

levels(mushrooms$odor)
levels(mushrooms$odor) <- c("almond","creosote","foul","anise","musty","none","pungent","spicy","fishy")

levels(mushrooms$habitat)
levels(mushrooms$habitat) <- c("woods","grasses","leaves","meadows","paths","urban","waste")
```

## Transformed Data
```{r pressure}
head(mushrooms)
```
<i>Users now have a better initial understanding of the data.</i>

## Summarize Data
```{r summarize-data}
summary(mushrooms)
```
